---
title: "p8105_hw1_dz2399"
author: "Dana Zeng"
date: "September 19, 2018"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

### Problem 1

##### Create a data frame comprised of:

* random sample of size 10 from a uniform[0, 5] distribution
* logical vector indicating whether elements of the sample are greater than 2
* (length-10) character vector
* (length-10) factor vector
```{r}
df = tibble(
  vec_numeric = runif(10, min=, max=5),
  vec_logic = (vec_numeric>2),
  vec_chr = c('a','b','c','a','b','c','a','b','c','a'),
  vec_factor = factor(c('a','b','c','a','b','c','a','b','c','a'))
)
df
```

##### Try to take the mean of each variable in your dataframe. What works and what doesn’t? Why?
```{r}
mean(df$vec_numeric)
mean(df$vec_logic)
mean(df$vec_chr)
mean(df$vec_factor)
```
The mean function works for numeric and logic value (taking true as 1 and false as 0), but it does not work for characters and factor values, returning NA. The reason is that in this case the factors are in characters, so r cannot make this factor and the character column to a numeric value, hence returning NA for means.

##### In some cases, you can explicitly convert variables from one type to another. Write a code chunk that applies the as.numeric function to the logical, character, and factor variables (please show this chunk but not the output). What happens? In a second code chunk, convert your character variable from character to factor to numeric; similarly, convert your factor variable from factor to character to numeric. What happens?
```{r, eval=F}
as.numeric(df$vec_logic)
as.numeric(df$vec_chr)
as.numeric(df$vec_factor)
```
##### What happened is that: 
* the logical value is turned into: true = 1, false = 0
* for the factor value, each level is automatically assigned value starting 1 to increasing integers
* the character value gives the NA

```{r}
as.numeric(as.factor(df$vec_chr))
as.numeric(as.character(df$vec_factor))
```

##### What happened is that:
* 'character to factor to numeric' returns numerical value, because characters can turned into factors, and then as.numeric assign numbers to each level of the factor.
* 'factor to character to numeric' returns NAs, because when factors are turned into characters, the as.numeric function will return NA for characters.


### Problem 2

##### Create a data frame comprised of:
* x: a random sample of size 1000 from a standard Normal distribution
* y: a random sample of size 1000 from a standard Normal distribution
* A logical vector indicating whether the x + y > 0
* A numeric vector created by coercing the above logical vector
* A factor vector created by coercing the above logical vector

```{r}
set.seed(1000)
df_plot = tibble(
  x = rnorm(1000),
  y = rnorm(1000),
  logic = x+y>0,
  as_numeric = as.numeric(logic),
  as_factor = as.factor(logic)
)
```
***
The sample size of the dataframe is `r dim(df_plot)[1]`, the mean of x is `r mean(df_plot$x)`, the median of x is `r median(df_plot$x)`, the proportion of cases for which the logical vector is TRUE is `r sum(df_plot$logic==T)/dim(df_plot)[1]`


##### Make a scatterplot of y vs x; color points using the logical variable (adding color = ... inside of aes in your ggplot code should help). Make a second and third scatterplot that color points using the numeric and factor variables, respectively, and comment on the color scales.

```{r}
ggplot(df_plot, aes(x = x, y = y,color=logic)) + 
  geom_point() + ggtitle('Color by logic')
ggplot(df_plot, aes(x = x, y = y,color=as_numeric)) + 
  geom_point() + ggtitle('Color by as_numeric')
ggplot(df_plot, aes(x = x, y = y,color=as_factor)) + 
  geom_point() + ggtitle('Color by as_factor')

ggsave('plot1_color_by_logic.jpg')

```

